### 2013-12-07 manuscripts/reply/reply-si-main.tex
Some minor editing of the reply ms.
Sending this version to co-authors.

%CWL:
%It may be important to point out the error made in their paper, but it can
%also be nit-picky, it depends on how the error impacts the overall message of
%our response.  We should strive to keep the response short and directed at the
%big picture items. If there is a way to still point out the error, then maybe
%some of this section should be used
\section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
\citet{Hickerson2013} re-analyzed the dataset of \citet{Oaks2012} using a
model-averaging approach, where they placed a discrete uniform prior over eight
different prior models (see Table 1 of \citet{Hickerson2013}).
However, there is a fundamental error in the methodology of
\citet{Hickerson2013}: their model mixes different time units.

@@

\section*{Differing utilities of \numt{} and \vmratio{} in \msb}
The main parameter of the \msb model is the vector of divergence
times for each of the taxon pairs,
$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
\citep{Oaks2012}.
\citet{Hickerson2013} argue that the dispersion index of this vector,
\vmratio{}, is a better model-choice estimator than the number of 
divergence time parameters within the vector,
\numt{}.
They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
\citet{Hickerson2013}), which is simply a plot of sample size versus variance.
This plot shows, not surprisingly, that \vmratio{} has very little information
about the number of divergences among taxa.
Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
biogeographically relevant than \numt{}.
% We struggle to follow this logic.
Certainly the maximum information is contained within the divergence time
vector that \vmratio{} is summarizing.
Clearly, the number of divergence time parameters within the vector and their
values is more informative than its variance (i.e., the dispersion index is not
a sufficient statistic for \divtvector).

\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
better than \numt{}.''
However, \msb is a model-choice estimator, and hence the goal is to estimate the
posterior probability of divergence models.
\citet{Oaks2012} demonstrate that even when all assumptions of the model are
met, \vmratio is a poor model-choice estimator (see plots B, D \& F of Figure
4), whereas \numt{} performs better.
Furthermore, \vmratio{} is limited to estimating the probability of only a
single divergence model (the one divergence model), and thus its utility for
model choice is extremely limited.

>>>>

\section*{An error in Hickerson et al.'s re-analysis of the Philippines data}
\citet{Hickerson2013} re-analyzed the dataset of \citet{Oaks2012} using a
model-averaging approach, where they placed a discrete uniform prior over eight
different prior models (see Table 1 of \citet{Hickerson2013}).
However, there was an error in their methodology; their model mixes different
units of time.

@@

\section*{Additional clarifications from \citet{Hickerson2013}}

\subsection*{Saturation of summary statistics}
\citet{Hickerson2013} claim the priors used by \citet{Oaks2012} ``cause much of
the explored parameter space to be beyond the threshold of saturation in most
mtDNA genes.'' To explore this possibility, we simulated datasets under prior
settings that match two of the three priors used by \citet{Oaks2012}:
$\meanDescendantTheta{} \sim U(0.0005, 0.04)$ and $\ancestralTheta{} \sim
U(0.0005, 0.02)$.
Under this prior, we draw divergence time parameters from a uniform
distribution of $U(0, 20)$ coalescent units, simulate datasets, and plot the
\divt{} values against the summary statistics calculated from the resulting
datasets (Figure~\ref{figSaturationPlot}).
Clearly, the priors used by \citet{Oaks2012} with upper limits on \divt{} of five
and 10 coalescent units suffered from little affect from saturation.
Even at divergence times of 20 coalescent units, there is still signal in the
summary statistics used by \msb (Figure~\ref{figSaturationPlot}).
Thus, the assertion of \citet{Hickerson2013} does not apply to at least
two of the priors used by \citet{Oaks2012} and, as a result, does not
explain the bias they found.

\subsection*{Graphical prior comparisons}
\citet{Hickerson2013} advocate the use of what they call graphical checks of
prior models.
This prior-predictive approach generating a small number (1000) of random
samples from the prior and plotting the resulting summary statistics in
comparison to the observed statistics to see if they coincide (see Figure 1 of
\citet{Hickerson2013}).
% As we show above, this strategy can be misleading, because the resulting plots
% of this approach have little correlation with the appropriateness of priors.
Given the richness of the \msb model ($\approx 600$ parameters for the Philippine
dataset analyzed by \citet{Hickerson2013}), we do not expect that 1000
\emph{random} draws from the vast prior parameter space will yield data and
summary statistics consistent with the observed data.
In fact, when such random draws are tightly clustered around the observed
statistics, this can be an indication that the prior is over-fit, as we show in
the main text (Table~\ref{tabModelChoiceEmpirical} and Figure~S\ref{figPCA}).
Thus, using such plots to select priors should be avoided, and the use of
posterior-predictive analyses would be much more informative about the overall
fit of models.

\subsection*{Differing utilities of \numt{} and \vmratio{} in \msb}
The main parameter of the \msb model is the vector of divergence
times for each of the taxon pairs,
$\divtvector = \{\divt{1}, \ldots, \divt{Y}\}$
\citep{Oaks2012}.
\citet{Hickerson2013} argue that the dispersion index of this vector,
\vmratio{}, is a better model-choice estimator than the number of 
divergence time parameters within the vector,
\numt{}.
They present a plot of \numt{} against \vmratio{} (Fig.~S1 of
\citet{Hickerson2013}), which is essentially a plot of sample size versus
variance.
This plot shows, not surprisingly, that \vmratio{} has very little information
about the number of divergences among taxa.
Nonetheless, \citet{Hickerson2013} conclude \vmratio{} more informative and
biogeographically relevant than \numt{}.
% We struggle to follow this logic.
Certainly the maximum information is contained within the divergence time
vector that \vmratio{} is summarizing.
Clearly, the number of divergence time parameters within the vector and their
values is more informative than its variance (i.e., the dispersion index is not
a sufficient statistic for \divtvector).

\citet{Hickerson2013} also argue that ``\msb can estimate \vmratio{} much
better than \numt{}.''
However, \citet{Oaks2012} demonstrate that even when all assumptions of the
model are met, \vmratio is a poor model-choice estimator (see plots B, D \& F
of Figure 4), whereas \numt{} performs better.
Furthermore, \vmratio{} is limited to estimating the probability of only a
single model (the one-divergence model), and thus its utility for
model choice is extremely limited.

### 2013-12-07 manuscripts/reply/short-reply.tex
Some minor editing of the reply ms.
Sending this version to co-authors.

As discussed by \citet{Oaks2012}, a straightforward prediction if marginal
likelihoods are causing the preference for simple models is that the bias should
disappear as the model generating the data converges to the prior.
\citet{Oaks2012} tested this prediction by performing 100,000 simulations to
assess the model-choice behavior of \msb when the prior model is correct.
The results confirm the prediction as \msb actually tends to underestimate
the probability of the one-divergence model (see Figure 4 of
\citet{Oaks2012}).
We confirmed this same behavior for the model-averaging approach used by
\citet{Hickerson2013} (see SI text and Figure~\ref{figValidationMCBehavior}).
These results are not clearly predicted if insufficient sampling was causing
the bias.
Even when the prior is correct, due to the discrete uniform prior on
\numt{} implemented in \msb, models with larger numbers of divergence
events (and thus greater parameter space) will still be less densely
sampled than those with fewer divergence events \citep{Oaks2012}.
Thus, the results of the simulations of \citet{Oaks2012} are more consistent
with the fundamental sensitivity of marginal likelihoods to priors.

@@

The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
of \msb are similar to those of \citet{Hickerson2013}, but avoid the
need for imposing an additional dimension of model choice.
\citet{Oaks2012} suggest that uniform priors may not be appropriate for many
parameters of the \msb model, and recommend the use of probability
distributions from the exponential family.
If we look at the prior distribution on divergence-time parameters imposed by
the model-averaging approach of \citet{Hickerson2013} we see it is a mixture of
overlapping uniforms with lower limits of zero (Figure~\ref{figMCTauPrior}).
This looks very much like an exponential distribution, except that in any state
of the model, all the divergence-time parameters are restricted to the hard
bounds of one of the uniform distributions.
Thus, it seems more appropriate to simply place a gamma prior (the exponential
being a special case of the gamma) on divergence times.
This would capture the prior uncertainty that \citet{Hickerson2013} are
suggesting for divergence times (Figure~\ref{figMCTauPrior}) while avoiding
costly model-averaging and the constraint that all divergence times must fall
within the hard bounds of the current model state.
It also would allow an investigator to place the majority of the prior density
in regions of parameter space they believe, \emph{a priori}, are most
plausible, but still capture uncertainty in the tails of distributions with low
density.
Most importantly, we have found the use of more flexible prior distributions in
place of uniforms improves the power of the method to detect temporal variation
in divergences and reduces the bias toward models of clustered divergences
\citep{Oaks2014dpp}.
% This seems like a natural solution to the problem of skewed marginal
% likelihoods and/or sampling error, which are both likely exacerbated by the
% fact that uniform distributions are the only choice of prior for many of the
% parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
% \descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
% \bottleScalar{1}{}, \bottleScalar{2}{}, \locusRateHetShapeParameter,
% \migrationRate{}, \recombinationRate;][]{Oaks2012}.

@@

\section*{Conclusions}
We demonstrate how the approximate Bayesian model choice method implemented in
\msb can be strongly biased away from models with greater parameter space.
As suggested by \citet{Oaks2012}, this is likely caused by the use of uniform
priors on most of the model's parameters.
Uniform distributions necessitate the use of broad priors that place high prior
density in unlikely regions of parameter space, less the risk of excluding the
truth \emph{a priori}.
These broad priors reduce the marginal likelihoods of models with greater parameter
space, either due to more divergence time parameters or broader prior
distributions on those parameters.
We show that the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} does not mitigate this bias, but rather causes it to
manifest by sampling predominantly from models that may exclude the true values
of the parameters.

>>>>

As discussed by \citet{Oaks2012}, a straightforward prediction if marginal
likelihoods are causing the preference for simple models is that the bias should
disappear as the model generating the data converges to the prior.
\citet{Oaks2012} tested this prediction by performing 100,000 simulations to
assess the model-choice behavior of \msb when the prior model is correct.
The results confirm the prediction as \msb actually tends to underestimate
the probability of the one-divergence model (see Figure 4 of
\citet{Oaks2012}).
We confirmed this same behavior for the model-averaging approach used by
\citet{Hickerson2013} (see SI text and Figure~S\ref{figValidationMCBehavior}).
These results are not clearly predicted if insufficient sampling was causing
the bias.
Even when the prior is correct, due to the discrete uniform prior on
\numt{} implemented in \msb, models with larger numbers of divergence
events (and thus greater parameter space) will still be less densely
sampled than those with fewer divergence events \citep{Oaks2012}.
Thus, the results of the simulations of \citet{Oaks2012} are more consistent
with the fundamental sensitivity of marginal likelihoods to priors.

@@

The recommendations of \citet{Oaks2012} for mitigating the lack of robustness
of \msb are similar to those of \citet{Hickerson2013}, but avoid the
need for imposing an additional dimension of model choice.
\citet{Oaks2012} suggest that uniform priors may not be appropriate for many
parameters of the \msb model, and recommend the use of probability
distributions from the exponential family.
If we look at the prior distribution on divergence-time parameters imposed by
the model-averaging approach of \citet{Hickerson2013} we see it is a mixture of
overlapping uniforms with lower limits of zero (Figure~\ref{figMCTauPrior}).
This looks very much like an exponential distribution, except that in any state
of the model, all the divergence-time parameters are restricted to the hard
bounds of one of the uniform distributions.
Thus, it seems more appropriate to simply place a gamma prior (the exponential
being a special case of the gamma) on divergence times.
This would capture the prior uncertainty that \citet{Hickerson2013} are
suggesting for divergence times (Figure~\ref{figMCTauPrior}) while avoiding
costly model-averaging and the constraint that all divergence times must fall
within the hard bounds of the current model state.
It also would allow an investigator to place the majority of the prior density
in regions of parameter space they believe, \emph{a priori}, are most
plausible, but still capture uncertainty in the tails of distributions with low
density.
Most importantly, we have found the use of more flexible distributions in place
of uniform priors improves the power of the method to detect temporal variation
in divergences and reduces the bias toward models of clustered divergences
\citep{Oaks2014dpp}.
% This seems like a natural solution to the problem of skewed marginal
% likelihoods and/or sampling error, which are both likely exacerbated by the
% fact that uniform distributions are the only choice of prior for many of the
% parameters of the \msb model \citep[\divt{}, \ancestralTheta{},
% \descendantTheta{1}{}, \descendantTheta{2}{}, \bottleTime{},
% \bottleScalar{1}{}, \bottleScalar{2}{}, \locusRateHetShapeParameter,
% \migrationRate{}, \recombinationRate;][]{Oaks2012}.

@@

\section*{Conclusions}
We demonstrate how the approximate Bayesian model-choice method implemented in
\msb can be strongly biased away from models with greater parameter space.
As suggested by \citet{Oaks2012}, this is likely caused by the use of uniform
priors on most of the model's parameters.
Uniform distributions necessitate the use of broad priors that place high prior
density in unlikely regions of parameter space, less the risk of excluding the
truth \emph{a priori}.
These broad priors reduce the marginal likelihoods of models with more
divergence-time parameters or broader prior distributions on those parameters.
We show that the empirical Bayesian model-averaging approach of
\citet{Hickerson2013} does not mitigate this bias, but rather causes it to
manifest by sampling predominantly from models that may exclude the true values
of the parameters.
